---
title: Urinary excretion of vitamin D binidng protein across varying degrees of kidney
  dysfunction and dysglycemia
author: "WINDY WANG"
date: "July 6, 2016"
# bibliography: bibliography.bib
output:
 word_document
    # reference_docx: ../docx_template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = '../')
```

```{r package}
library(nephro)
library(carpenter)
library(ggplot2)
library(knitr)
library(plyr)
library(dplyr)
library(tidyr)
library(pander)
library(captioner)
library(knitr)
library(mason)
```

```{r data}
ds <- readRDS(file='data/ds.Rds')
source('R/functions.r')
```

# Values
```{r eGFR}
ds_base <- ds %>% 
  filter(VN == 1, UDBP < 5000)

ds %>% 
  filter(VN == 1, UDBP < 5000) %>% 
  select(dm_status, udbpCrRatio) %>% 
  na.omit() %>% 
  group_by(dm_status) %>% 
  summarise(Median = median(log(udbpCrRatio)),
            Lower = quantile(log(udbpCrRatio), probs = 0.25),
            Upper = quantile(log(udbpCrRatio), probs = 0.75)) %>% 
  knitr::kable()
  
anova <- aov(log(ds_base$udbpCrRatio) ~ ds_base$mcr_status)
summary(anova)
TukeyHSD(anova)

rm(anova)
rm(ds_base)
```

find n
```{r n}
table(ds$fVN)
```


# Subject Characteristics
```{r sub-char}
ds %>% 
  filter(VN == 1) %>% 
  table_baseline('UDBP_status', "Subject characteristics")
  
# ANOVA for subject characteristics table
anova <- aov(ds$Glucose120~ds$fVN)
summary(anova)
TukeyHSD(anova)

rm(anova)

# Chi-square
chisq.test(table(ds$dm_status, ds$fVN), correct = TRUE)
```


# Graphs

Scatterplot: UDBP adjustment
```{r scatter-UDBP}
ds2 <- ds %>% 
  filter(VN == 1)

ds2 %>% 
  scatter_plot("UDBP", "udbpCrRatio",
               "Urinary VDBP (ng/mL)", "uVDBP:creatinine") +
  geom_smooth(colour = "grey50")

###################

cor.test(ds2$UDBP, ds2$udbpCrRatio, method = "spearman", exact = FALSE)

rm(ds2)
```

Scatterplot: uVDBP and uCreatinine
```{r scatter-creat}
ds %>% 
  filter(VN == 1) %>% 
  scatter_plot("UrineCreatinine", "UDBP", 
               "Urinary Creatinine (mmol/L)", "Urinary VDBP (ng/mL)") +
  geom_smooth(colour = "grey50")
```

Histogram
```{r histogram}
ds %>% 
  filter(VN == 1) %>% 
  histo_plot("UDBP", 50, "Urinary VDBP (ng/mL)")
```

Histogram: Longitudinal
```{r histogram2}
ds_visit <- ds %>% 
  filter(UDBP < 5000) %>% 
  select(fVN, UDBP, UDBP_status) %>% 
  na.omit() %>% 
  group_by(fVN) %>% 
  mutate(n = ifelse(row_number()==n(), paste("n =", n()), NA))

ds_visit %>% 
  histo_plot("log(UDBP)", 1, "log urinary VDBP (ng/mL)") +
  facet_wrap(~fVN, switch = "x") +
  geom_text(aes(label = n), y = 230, size = 3, colour = "grey40")
```


Boxplot: eGFR categories
```{r box-eGFR}
ds_base <- ds %>% 
  filter(VN == 1)

ds_base %>% 
  select(eGFR_status, udbpCrRatio, UDBP) %>% 
  na.omit() %>% 
  box_plot2("eGFR_status", "log(udbpCrRatio)", 
            "Estimated GFR (ml/min/1.73m^2)",
            "log uVDBP:Creatinine") +
  scale_x_discrete(labels = paste(levels(ds_base$eGFR_status),
                                    "\n(n=", table(ds_base$eGFR_status), ")", sep = ""))

anova <- aov(log(ds_base$UDBP)~ds_base$eGFR_status)
summary(anova)
# TukeyHSD(anova)

rm(ds_base)
rm(anova)
```

Longitudinal boxplot: eGFR categories
```{r box-eGFR2}
ds_visit <- ds %>% 
  filter(UDBP < 5000) %>% 
  select(fVN, UDBP, udbpCrRatio, eGFR_status) %>% 
  na.omit() %>% 
  group_by(fVN, eGFR_status) %>% 
  mutate(n = ifelse(row_number()==n(), paste("n =", n()), NA))

ds_visit %>% 
  box_plot2("eGFR_status", "log(UDBP)",
            "Estimated GFR Status", "log UDBP") +
  theme(axis.title.x = element_blank()) +
  facet_grid(~fVN, switch = "x") +
  geom_text(aes(label = n), y = -7 , size = 3, colour = "grey40")

###########################

ds_visit <- ds %>% 
  filter(UDBP < 5000, VN == 6)

anova <- aov(log(ds_visit$UDBP)~ds_visit$eGFR_status)
summary(anova)
# TukeyHSD(anova)

rm(anova)
rm(ds_visit)
```


Scatterplot: eGFR 
```{r scatter-eGFR}
ds_base <- ds %>% 
  filter(VN == 1, UDBP < 5000)

ds_base %>% 
  select(eGFR, UDBP, UrineCreatinine) %>% 
  na.omit() %>% 
  scatter_plot("eGFR", "log(UDBP/UrineCreatinine)",
               "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
               "log UDBP:Creatinine") +
  geom_smooth(method = lm, colour = "grey50")

cor.test(ds_base$eGFR, ds_base$udbpCrRatio, method = "spearman", exact = FALSE)

rm(ds_base)
```

Longitudinal scatterplot: eGFR
```{r scatter-eGFR2}
ds %>% 
  filter(UDBP < 5000) %>% 
  scatter_plot("eGFR", "log(UDBP)",
            "Estimated GFR (ml/min/1.73m^2)", "log uVDBP") +
  geom_smooth(colour = "grey50") +
  facet_grid(~fVN, switch = "x")

####################

ds_visit <- ds %>% 
  filter(VN == 6, UDBP < 5000)

cor.test(ds_visit$eGFR, ds_visit$UDBP, method = "spearman", exact = FALSE)

rm(ds_visit)
```


Boxplot: MCR categories
```{r box-mcr}
ds_mcr <- ds %>% 
  mutate(mcr_cat = ifelse(mcr_status == "Normal", "Normal", "Albuminuria")) %>% 
  mutate(mcr_cat = factor(mcr_cat,
                          levels = c("Normal", "Albuminuria"),
                          ordered = TRUE)) %>% 
  filter(VN == 1, UDBP < 5000)

ds_mcr %>% 
  filter(UDBP < 5000, CRP < 10) %>% 
  select(mcr_status, udbpCrRatio) %>% 
  na.omit() %>% 
  box_plot2("mcr_status", "log(udbpCrRatio)", 
            "Albuminuria",
            "log uVDBP:Creatinine") +
  scale_x_discrete(labels = paste(levels(ds_mcr$mcr_status), 
                                    "\n(n=", table(ds_mcr$mcr_status), ")", sep = ""))

anova <- aov(ds_mcr$UDBP~ds_mcr$mcr_status)
summary(anova)
TukeyHSD(anova)

rm(ds_mcr)
rm(anova)
```

Longitudinal boxplot: MCR categories
```{r box-mcr2}
ds_visit <- ds %>% 
  mutate(mcr_cat = ifelse(mcr_status == "Normal", "Normal", "Albuminuria")) %>% 
  mutate(mcr_cat = factor(mcr_cat,
                          levels = c("Normal", "Albuminuria"),
                          ordered = TRUE)) %>% 
  select(fVN, UDBP, udbpCrRatio, mcr_cat) %>% 
  na.omit() %>% 
  group_by(fVN, mcr_cat) %>% 
  mutate(n = ifelse(row_number()==n(), paste("n =", n()), NA))

ds_visit %>% 
  box_plot2("mcr_cat", "log(udbpCrRatio)",
            "Estimated GFR Status", "log UDBP:Creatinine") +
  theme(axis.title.x = element_blank()) +
  facet_grid(~fVN, switch = "x") +
  geom_text(aes(label = n), y = 5 , size = 3, colour = "grey40")

rm(ds_visit)
```

Scatterplot: MCR categories
```{r scatter-mcr}
ds_base <- ds %>% 
  filter(VN == 1, UDBP < 5000)

ds_base %>% 
  select(MicroalbCreatRatio, UDBP, UrineCreatinine) %>% 
  na.omit() %>% 
  scatter_plot("log(MicroalbCreatRatio)", "log(UDBP/UrineCreatinine)",
               "log Microalbumin:Creatinine Ratio",
               "log UDBP:Creatinine") +
  geom_smooth(method = lm, colour = "grey50")

cor.test(ds_base$MicroalbCreatRatio, ds_base$udbpCrRatio, 
         method = "spearman", exact = FALSE)

rm(ds_base)
```

Longitudinal scatterplot: MCR
```{r scatter-mcr2}
ds %>% 
  filter(UDBP < 5000) %>% 
  scatter_plot("log(MicroalbCreatRatio)", "log(UDBP/UrineCreatinine)",
            "log Microalbumin-to-Creatinine Ratio", "log uVDBP:Creatinine") +
  geom_smooth(colour = "black") +
  facet_grid(~fVN, switch = "x")
```

Boxplot: DM categories
```{r box-DM}
ds_base <- ds %>% 
  filter(VN == 1, UDBP < 5000)

ds_base %>% 
  select(dm_status, UDBP, UrineCreatinine) %>% 
  na.omit() %>% 
  box_plot2("dm_status", "log(UDBP/UrineCreatinine)", 
            "Glycemic Status",
            "log uVDBP:Creatinine") +
  scale_x_discrete(labels = paste(levels(ds_base$dm_status), 
                                  "\n(n=", table(ds_base$dm_status), ")", sep = ""))

anova <- aov(ds_base$udbpCrRatio~ds_base$dm_status)
summary(anova)
# TukeyHSD(anova)

# kruskal.test(ds_base$UDBP ~ ds_base$mcr_status) 

rm(anova)
rm(ds_base)
```

Scatterplot: Fasting glucose
```{r scatter-fasting}
ds %>% 
  filter(VN == 1) %>% 
  filter(UDBP < 5000) %>% 
  select(Glucose0, UDBP, UrineCreatinine) %>% 
  na.omit() %>% 
  scatter_plot("Glucose0", "log(UDBP/UrineCreatinine)",
               "Fasting Glucose (mmol/L)",
               "log UDBP:Creatinine") +
  geom_smooth(colour = "black")
```

Scatterplot: 2h OGTT
```{r scatter-2h}
ds %>% 
  filter(VN == 1) %>% 
  filter(UDBP < 5000) %>% 
  select(Glucose120, UDBP, UrineCreatinine) %>% 
  na.omit() %>% 
  scatter_plot("Glucose120", "log(UDBP/UrineCreatinine)",
               "2h Post-OGTT Blood Glucose (mmol/L)",
               "log UDBP:Creatinine") +
  geom_smooth(colour = "black")
```

Scatterplot: 25(OH)D
```{r scatter-vitD}
ds_base <- ds %>% 
  filter(VN == 1, UDBP < 5000)

ds_base %>% 
  scatter_plot("log(udbpCrRatio)", "VitaminD",
               "log uVDBP:creatinine",
               "Serum 25(OH)D") +
  geom_smooth(method = lm, colour = "grey50")

cor.test(ds_base$VitaminD, ds_base$UDBP, method = "spearman", exact = FALSE)

rm(ds_base)
```

Subgroup analysis: 25(OH)D and UDBP
```{r subgroup}
# Vitamin D status - Endocrine Society:
#     Deficient = <50 nmol/L
#     Insufficient = 50-74
#     Sufficient = >74

# Determine PTH tertiles (<3.61, 3.61 - 5.0, >5.0)
# ds %>% 
#   filter(VN == 1, UDBP < 5000) %>% 
#   summarise(Lower = quantile(PTH, probs = 0.33, na.rm = TRUE),
#             Upper = quantile(PTH, probs = 0.67, na.rm = TRUE))

# Calculate n
# table(ds_sub$fMedsBP)

ds_sub <- ds %>% 
  filter(VN == 1, UDBP < 5000) %>% 
  mutate(dm_status2 = ifelse(dm_status == "NGT", "NGT",
                              "Dysglycaemia"),
         mcr_status2 = ifelse(mcr_status == "Normal", "Normal",
                              "Albuminuria"),
         eGFR_status2 = ifelse(eGFR_status == "Normal", "Normal", 
                               "KD"),
         PTH_status = ifelse(PTH < 3.61, "PTH1",
                             ifelse(PTH > 5, "PTH3",
                                    "PTH2")),
         VitD_status = ifelse(VitaminD < 50, "Deficient", 
                              ifelse(VitaminD > 74, "Sufficient", 
                                     "Insufficient")),
         ALT_status = ifelse(ALT > 6 & ALT < 57, "Normal",
                             "Abnormal")) %>% 
  # group_by(mcr_status) %>%
  # mutate(n = ifelse(row_number()==n(), paste("n=", n()), NA))
  filter(Sex == "Female")

ds_sub %>%
  select(udbpCrRatio, VitaminD, mcr_status) %>%
  na.omit() %>%
  scatter_plot("log(udbpCrRatio)", "VitaminD",
               "log UDBP:Creatinine",
               "Serum 25(OH)D (nmol/L)") +
  geom_smooth(method = "lm", colour = "grey50") 
  # facet_grid(~mcr_status, switch = "x") 
  # theme(strip.text.x = element_text(size = 14, colour = "grey70"))
  # geom_text(aes(label = n), y = 0, size = 12, colour = "grey40")

cor.test(ds_sub$VitaminD, ds_sub$UDBP, method='pearson', exact=FALSE)

rm(ds_sub)
```

Longitudinal look at 25(OH)D and uVDBP
```{r scatter-vitD2}
ds %>% 
  filter(UDBP < 5000) %>% 
  scatter_plot("log(UDBP/UrineCreatinine)", "VitaminD",
            "log uVDBP:Creatinine", "Serum 25(OH)D") +
  geom_smooth(method = "lm", colour = "grey50") +
  facet_grid(~fVN, switch = "x")
```


Multivariate Analysis
```{r ancova}
ds2 <- ds %>% 
  filter(VN == 1) %>% 
  select(VitaminD, UDBP, Age, Sex, Ethnicity, BMI, dm_status, eGFR_status, udbpCrRatio) %>% 
  na.omit()

model1 <- lm(VitaminD ~ udbpCrRatio, 
             data = ds2)
model2 <- lm(VitaminD ~ udbpCrRatio + Age + Sex + Ethnicity + BMI, 
             data = ds2)
model3 <- lm(VitaminD ~ udbpCrRatio + Age + Sex + Ethnicity + BMI + dm_status, 
             data = ds2)
model4 <- lm(VitaminD ~ udbpCrRatio + Age + Sex + Ethnicity + BMI + dm_status + eGFR_status, 
             data = ds2)

anova(model1, model2, model3, model4)
AIC(model1, model2, model3, model4)
summary(fit)



# independence = assumes independence between clusters (same as linear regression)
# autoregressive = correlation between the two variables gets smaller as years pass
# exchangable = correlation doesn't change between time points
# unstructured = cannot predict how correlation changes between time points 
# GEE calculates the working correlation, not the actual 

multivar <- ds %>% 
  dplyr::filter(VN == 1, UDBP < 5000) %>% 
  design("glm") %>% 
  add_settings(family = gaussian()) %>% 
  add_variables("yvars", "VitaminD") %>% 
  add_variables("xvars", "udbpCrRatio") %>%
  construct() %>% 
  add_variables("covariates", c("Age", "Sex", "Ethnicity", "BMI",
                                "dm_status",
                                "eGFR_status")) %>%
  construct() %>% 
  scrub()

  add_variables("covariates", "dm_status") %>%
  construct() %>%
  add_variables("covariates", "eGFR_status") %>%
  construct() %>%
  add_variables("covariates", "mcr_status") %>%
  construct() %>%
  add_variables("covariates", "fMedsBP") %>%
  construct() %>% 
  scrub()

# PROMISE_data %>% 
#   design("gee") %>% 
#   add_settings(cluster.id = "SID", 
#                corstr = "ar1",
#                family = gaussian()) %>% 
#   add_variables("yvars", "ISI") %>% 
#   add_variables("xvars", c("BMI", "Weight")) %>% 
#   add_variables("covariates", c("Sex", "Ethnicity", "VN")) %>% 
#   construct() %>% 
#   add_variables("yvars", c("HOMA")) %>% 
#   add_variables("covariates", "VN") %>% 
#   construct() %>% 
#   scrub() %>% 
#   polish_filter("Xterm$", "term") %>% 
#   polish_adjust_pvalue() %>% 
#   polish_transform_estimates(function(x) exp(x)) # "Back-transform" if log
  
```

PTH and VitaminD
```{r 3D}
ds2 <- ds %>% 
  filter(VN == 1, UDBP < 5000) %>% 
  mutate(PTH_status = ifelse(PTH < 3.61, "PTH1",
                             ifelse(PTH > 5, "PTH3",
                                    "PTH2")),
         VitD_status = ifelse(VitaminD < 50, "Deficient", 
                              ifelse(VitaminD > 74, "Sufficient", 
                                     "Insufficient")),
         ALT_status = ifelse(ALT > 6 & ALT < 57, "Normal",
                             "Abnormal")) 

ds2 %>% 
  filter(VN == 1) %>% 
  ggplot(aes(PTH_status, log(udbpCrRatio))) +
  geom_point(mapping = aes(colour = VitD_status))

rm(ds_visit)
```


