---
title: "Serum 25(OH)D Relationships"
author: "WINDY WANG"
date: "May 13, 2016"
# bibliography: bibliography.bib
output: 
  word_document:
    reference_docx: docx_template.docx
---

```{r knit_setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(collapse = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE)

library(carpenter)
library(ggplot2)
library(knitr)
library(plyr)
library(dplyr)
library(tidyr)
library(pander)

ds_clean <- readRDS(file="ds_clean.Rds")
ds <- readRDS(file='ds.Rds')
source('R/functions.r')
```
## Purpose

The aim of this document is to explore how serum 25(OH)D levels relate to other variables. This exploration of the data may lead to better understanding of the interaction of urinary vitamin D binding protein and serum levels of vitamin D.

# Kidney Status

In a paper by Kalousova et al. (2015), it was found that plasma vitamin D levels decreased in subjects with worse kidney function (i.e. chronic kidney disease and long-term haemodialysis).

## Estimated Glomerular Filtration Rate

In the PROMISE cohort at baseline, it appears that there is a significant relationship between estimated glomerular filtration (eGFR) rate and serum levels of 25(OH)D. Although the Spearman's coefficient is negative (r = -0.19), a LOESS curve of the data shows a somewhat parabolic relationship. Subjects with eGFR classified as normal (90 - 125 ml/min/1.73m^2, according to National Kidney Foundation guidelines) tended to have higher levels of 25(OH)D, but subjects outside that range tended to have lower levels of serum 25(OH)D.

```{r eGFR}
cor.test(ds$eGFR, ds$VitaminD, method='spearman', exact=FALSE)

ds %>% 
  scatter_plot(ds$eGFR, ds$VitaminD, 
               'Estimated Glomerular Filtration Rate (ml/min/1.73m^2)', 
               'Serum 25(OH)D (nmol/L)') +
  geom_smooth(se=TRUE, colour='black')
```

## Microalbuminuria
```{r albumin}
cor.test(ds$MicroalbCreatRatio, ds$VitaminD, method='spearman', exact=FALSE)

box.plot(ds$mcr_status, ds$VitaminD, 
             'Microalbuminuria', 
             'Serum 25(OH)D (nmol/L)')

```

# PTH-Calcium Axis

```{r PTH-ratio}
scatter_plot(ds$PTH, ds$CaCrRatio,
             'PTH', 'Urinary Calcium:Creatinine') + 
  geom_smooth(method=lm, se=TRUE, colour='black')
```

```{r}
scatter_plot(ds$UDBP_cr_ln, ds$CaCrRatio,
             'log UDBP:Creatinine', 'Urinary Calcium:Creatinine') +
  geom_smooth(colour='black')
```


# Diabetic Status

In accordance with previous literature, subjects who have diabetes (DM) tend to have a worse vitamin D status (*p*=0.005). It was also observed that subjects who are classified as prediabetic also have low 25(OH)D levels. No significant differences were found between 25(OH)D concentrations in those with prediabetes and normal glucose tolerance (NGT) (*p*=0.78) likely due to the smaller sample size in the prediabetes population (n=38), but there was a significant difference observed between subjects with prediabetes and diabetes (*p*=0.04), as well as NGT and diabetes (*p*=0.005).

```{r dm.status}
bar.plot(ds$dm_status, ds$VitaminD, 
         'Diabetic Status', 'Serum 25(OH)D (nmol/L)')

anova <- aov(ds$VitaminD~ds$dm_status)
summary(anova)
TukeyHSD(anova)
```

# Possible explanations for no association between UDBP and 25(OH)D

Urinary VDBP loss is normally compensated for by synthesis in the liver, which is estimated at 10 mg/kg [@Laing2005]
* Patients with liver disease was previously found to have lower plasma VDBP and 25(OH)D [@Bikle1986]
* If liver function is normal, VDBP levels may not change, so there will not be an association between VDBP and 25(OH)D

The occupancy of VDBP by vitamin D metabolites is <5% [@Speeckaert2006]
* Only severe loss of VDBP would result in 25(OH)D deficiency
* Other mechanisms likely plays a larger role
    + Decreased sun exposure
    + Disturbed conversion of vitamin D precursor on the skin
    + Decreased dietary intake
    + Accelerated vitamin D catabolism

The PROMISE cohort appears to be vitamin D deficient (mean 25(OH)D = 55.4 nmol/L; >75 nmol/L indicates sufficient vitamin D stores [@Kalousova2015])
