---
title: Preliminary analysis of estimated glomerular filtration rate using the PROMISE
  cohort at baseline
author: "WINDY WANG"
date: "May 12, 2016"
output:
  word_document:
    reference_docx: ../docx_template.docx
---

```{r knit_setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')
knitr::opts_chunk$set(collapse = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE)

library(carpenter)
library(ggplot2)
library(knitr)
library(plyr)
library(dplyr)
library(tidyr)
library(captioner)
library(knitr)

tbls <- captioner(prefix = 'TABLE')
figs <- captioner(prefix = 'FIGURE')

ds <- readRDS(file='ds.Rds')
ds_base <- ds %>% 
  filter(VN == 1)
source('R/functions.r')
```

```{r table_label}
tbl_char1 <- tbls(name='sub-char1',
                 caption="Subject characteristics according to estimated GFR categories at baseline.")

tbl_char3 <- tbls(name='sub-char3',
                 caption="Subject characteristics according to estimated GFR categories at 3-year visit.")

tbl_char6 <- tbls(name='sub-char6',
                  caption = "Subject characteristics according to estimated GFR categories at 6-year visits.")
```

# Subject Characterization

```{r sub-char1}
ds %>% 
  filter(VN == 1) %>% 
  table_baseline ('eGFR_status', tbl_char1)
```

```{r sub-char3}
ds %>% 
  filter(VN == 3) %>% 
  table_baseline('eGFR_status', tbl_char3)
```

```{r sub-char6}
ds %>% 
  filter(VN == 6) %>% 
  table_baseline('eGFR_status', tbl_char6)
```

## Moderate to Severe eGFR
### Cross-sectional at Baseline

At baseline, there were `r ds_base %>% filter(eGFR_status == 'Moderate') %>% summarise(n=n())` people who had estimated glomerular filtration rate (eGFR) of less than 60 ml/min/1.73m^2. These individuals are classified as having moderate kidney dysfunction according to the National Kidney Foundation. Upon taking a closer look at these individuals, their eGFR values are only slightly below the 60 ml/min/1.73m^2^ cut-off. These two subjects had missing values for both 3 year and 6 year visits.

```{r}
ds_base %>% 
    filter(SID %in% c(2075, 2266)) %>% 
    mutate(Baseline = round(eGFR, 1)) %>% 
    select(SID, Baseline) %>% 
    knitr::kable()
```

### Prospective
At the 3 year follow-up visit, there were `r ds %>% filter(VN==3, eGFR_status=='Moderate') %>% summarise(n=n())` subjects who had eGFR less than 60 ml/min/1.73m^2. This number decreased to `r ds %>% filter(VN==6, eGFR_status=='Moderate') %>% summarise(n=n())` at the 6 year follow-up. The lowest eGFR was `r min(ds$eGFR) %>% round(1)`ml/min. There were two subjects with eGFR of <45 ml/min/1.73m^2, which is classified as moderate to severe kidney dysfunction. Unfortunately, no subjects have eGFR values across all time-points, making progression of the disease difficult to analyse. There were two subjects with eGFR measurements at the 3 year and 6 year time points. Their eGFR either did not change much (`r ds %>% filter(VN==3, SID==1240) %>% select(eGFR) %>% round(1)`ml/min at 3 year and `r ds %>% filter(VN==6, SID==1240) %>% select(eGFR) %>% round(1)`ml/min at 6 year) or decreased (`r ds %>% filter(VN==3, SID==1295) %>% select(eGFR) %>% round(1)`ml/min at 3 year and `r ds %>% filter(VN==6, SID==1295) %>% select(eGFR) %>% round(1)`ml/min at 6 year).

Below is a table of subjects who had eGFR < 60 ml/min across all time points.

```{r}
SID_LowGFR <- ds %>% 
    filter(eGFR < 60) %>% 
    select(SID) %>% 
    .[[1]]

ds %>% 
    filter(SID %in% SID_LowGFR) %>% 
    select(SID, VN, eGFR, Creatinine, UDBP, dm_status, mcr_status) %>% 
    arrange(SID, VN) %>% 
    knitr::kable()
```

The two subjects with eGFR across multiple time points are emphasized in bold.

Only two subjects had eGFR < 50, so a more detailed analysis was conducted.

SID 3115 had a much lower eGFR compared to others, so I took a more detailed look at his/her eGFR across different time points. It may be that there was a problem with serum creatinine, as the eGFR is in the healthy range at baseline and 6-year.

Another subject (SID 1508) also had relatively low eGFR. Based on their trend, it appears that their kidney filtration deterioated as time passed. However, they had normal MCR and were not diabetic. His/Her urinary VDBP dropped drastically at visit 3, even though other parameters (e.g. eGFR, diabetic status, MCR) that we thought were associated with UDBP did not change much. At visit 6, UDBP increased to around 90 ng/mL again. 
