---
title: "Exploration of the relationship between parathyroid hormone and urinary vitamin D binding protein"
author: "WINDY WANG"
date: "May 25, 2016"
output: 
  word_document:
    reference_docx: docx_template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE)

library(ggplot2)
library(knitr)
library(plyr)
library(dplyr)
library(tidyr)
library(captioner)

tbls <- captioner(prefix = 'Table')
figs <- captioner(prefix = 'Figure')

ds <- readRDS(file='ds.Rds')
source('R/functions.r')
```

```{r figure_label}
fig_scatter <- figs(name='calcium-calcium:cr',
     caption="Association between urinary calcium and urinary calcium adjusted for urine volume using urinary creatinine. Significance was assessed using Spearman's Rank Correlation, r = 0.69, p < 0.001.")

fig_PTH <- figs(name='PTH-UCa',
     caption="Association between parathyroid hormone and urinary calcium. Significance was assessed using Spearman's Rank Correlation, r = -0.15, p < 0.001.")
```

## Urinary Calcium: Adjusted vs Unadjusted

```{r fig_scatter, fig.cap=fig_scatter}
ds %>% 
  scatter_plot('UrinaryCalcium', 'CaCrRatio',
             'Urinary Calcium (mmol/L)', 'Urinary Calcium : Urinary Creatinine') +
  geom_smooth(method=lm, colour='black')
```

Generally, there is a linear relationship between urinary calcium adjusted and unadjusted for urine volume (r = 0.69, *p* < 0.001), a few points deviate from the regression line. 

All subsequent analysis will be performed using the adjusted urinary calcium variable to reduce variation between subjects.


## Urinary calcium and Parathyroid Hormone

```{r PTH, fig.cap=fig_PTH}
ds %>% 
  scatter_plot('PTH', 'CaCrRatio',
             'Parathyroid Hormone (pmol/L)', 'Urinary Calcium:Creatinine') +
  geom_smooth(method=lm, colour='black')

cor_pvalue <- cor.test(ds$Diastolic, ds$UDBP_cr_ln, method='spearman', exact=FALSE)$p.value %>% 
  round(4)
```

There is a slight negative relationship between parathyroid hormone (PTH) and urinary calcium (r = `r cor(ds$PTH, ds$CaCrRatio, use="complete.obs") %>% round(2)`, p = `r cor_pvalue`). PTH increases renal reabsorption of calcium in order to increase serum levels. As expected, there is less urinary calcium when serum PTH concentrations are higher. 


## Serum 25(OH)D and Urine Calcium

```{r}
ds %>%
  scatter_plot('VitaminD', 'CaCrRatio',
             'Serum 25(OH)D (nmol/L)', 'Urinary Calcium:Creatinine') +
  geom_smooth(method=lm, colour='black')

figs(name="VitD-UCa",
     caption="Association between serum 25(OH)D and urinary calcium. Significance was assessed using Spearman's Rank Correlation, r = 0.07, p = 0.07.")
```

There is a slight non-significant positive relationship between 25(OH)D and urinary calcium (r = 0.07, *p* = 0.07), possibly reflecting the a positive relationship between 25(OH)D and serum calcium. The weak association is likely because two different types of biomarkers are compared, i.e. serum versus urinary.


<!-- ### Subgroup analysis -->

<!-- The PROMISE cohort at baseline is still relatively healthy, and there are only 8 subjects with urinary calcium levels in the hypercalciuria range (>7 mmol/L). However, the relationship between PTH and urinary calcium may be stronger in certain subgroups. -->

<!-- #### Diabetes -->

<!-- ```{r} -->
<!-- ds %>% -->
<!--   filter(DM == 1) %>% -->
<!--   scatter_plot("PTH", "CaCrRatio",  -->
<!--                "Parathyroid Hormone (pmol/L)", "Urinary Calcium:Creatinine") -->
<!-- ``` -->
<!-- The relationship between  -->




<!-- ## Serum 25(OH)D and urinary VDBP -->

<!-- Overall, there appears to be a very weak positive association between 25(OH)D and UDBP.  -->

<!-- ```{r vitD_UDBP} -->
<!-- scatter_plot(ds$VitaminD, ds$UDBP_cr_ln,  -->
<!--              'Serum 25(OH)D (nmol/L)', 'log UDBP:Creatinine') + -->
<!-- geom_smooth(se=TRUE, colour='black') -->
<!-- ``` -->

<!-- ## Parathyroid Hormone -->

<!-- Low vitamin D -> low calcium -> high PTH; -->
<!-- Therefore, we expect a negative correlation between vitamin D and PTH; -->
<!-- Consequently, if low serum 25(OH)D is correlated with higher urinary VDBP, we expect a positive association between UDBP and PTH. -->

<!-- ```{r PTH1} -->
<!-- scatter_plot(ds$PTH, ds$VitaminD,  -->
<!--              'Parathyroid hormone (pmol/L)', 'Serum 25(OH)D (nmol/L)') + -->
<!-- geom_smooth(se=TRUE, colour='black') # method=lm for linear line -->
<!-- ``` -->

<!-- ```{r PTH2} -->
<!-- scatter_plot(ds$PTH, ds$UDBP_cr_ln,  -->
<!--              'Parathyroid hormone (pmol/L)', 'log UDBP:Creatinine') + -->
<!-- geom_smooth(se=TRUE, colour='black') # method=lm for linear line -->
<!-- ``` -->

